### 2023-06-14 Conference Submission/DraftDoc.tex
Final Thesis Copy

\documentclass[12pt]{amsart}
\usepackage[utf8]{inputenc}
\usepackage[
backend=biber,
style=numeric,
maxbibnames=99,
sorting=ynt
]{biblatex}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{svg}
\usepackage{gensymb}
\usepackage{float}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx, animate}
\usepackage{caption, subcaption}
\usepackage{pgffor}
\usepackage{mathtools}
\usepackage{array}
\addbibresource{bibliography.bib}
\usepackage{hyperref,thmtools}
\linespread{1.5}

% Override ugly default link
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = blue    %Colour of citations
}

@@

\begin{section}{Introduction}
    Temporal networks are networks whose nodes and edges change in time. In this thesis we look to model temporal networks. That is, describe the changing temporal network using mathematical language. 
    % Why temp nets are important
    The modelling of temporal networks is an important task in many real world applications including symptom interactions for mental health \cite{jordan2020current,contreras2020temporal}, epidemiology \cite{masuda2013predicting}, and protein interactions \cite{lucas2021inferring,jin2009identifying}.

    % Temp nets are dynam systs
    Temporal networks can be seen as a system in which points, in our case nodes in a network, whose states, the edges connecting them, vary dependent in time, a dynamical system.

    % Hard to find dynam systs because of discrete events 
    Discovering the underlying equations governing these dynamical systems proves challenging. That is because changes in network structure are typically observed in the form of discrete jumps from one state to another; for example an edge between two nodes not being observed at the first time step then being observed at the next.

    % We are proposing a solution
    In this thesis, we propose a hybrid statistical and deep learning framework. That is a framework in which a neural network is used to model probabilistic values of the temporal network. This framework allows us to model temporal networks as continuous-time dynamical systems, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a temporal network.

    % DEs are useful for modelling interacting states
    Differential equations are useful for modelling systems where the state of one variable can effect the trajectories of other variables. We observe this behavior in temporal networks; nodes' connections within the network can influence the observation of edges between other nodes, for example the phenomenon observed in \cite{newman2001clustering,capocci2006preferential}, where a node is more likely to gain the connections the more connections it has (preferential attachment). With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling.

    % Temp net events are discrete
    In the common representation of networks as binary-valued adjacency matrices, matrices where the observation or lack thereof, of an edge is represented as a 1 or 0 respectively, the events recorded in a temporal sequence of networks correspond to the appearance or the disappearance of link.

    % Difficult to model discrete events
    Because of the discrete nature of events, directly modelling the temporal networks as dynamical systems would require us to handle discrete jumps. The discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

@@

    % Our solution is interpretable because we use symb reg
    We then define and use Neural Network Differential Equations (NNDE)\cite{SciML_C_Rak} to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow us to predict forward in time the evolution of the temporal networks.

@@

    % The framework can be applied to lots of fields
    As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations (UDEs).
\end{section}

\begin{section}{Literature Review}

    \subsection{Temporal Networks}
        % Use of temp nets
        \subsubsection{Definitions}
        Temporal networks are used to describe a series of interactions across time.
        A temporal network $G$ is defined as:  
        \begin{align}
            G=\{ {_t} G=({_t} N, {_t} E)\} _{t \in 1\ldots T } 
        \end{align}
        $_t x \in {_t} N, {_t} e\in {_t} E$, where ${_t} x, {_t} e$ are nodes and edges at time $t$ respectively, and ${_t} N, {_t} E$ are the sets of nodes and edges at time $t$ respectively.
        An example of this can be found in \autoref{method nets}.

        %representations of temp nets as adj lists
        This can be represented in many ways, for example as a series of interactions. The series of interactions is generally represented as a list of tuples containing the nodes which interacted and the time that the interaction took place. 
        \begin{align}
            G=\{(i, j, t)|\space i,j\in {_t} N, (i\leftrightarrow j)\in {_t}E \}
        \end{align}

@@

        \subsubsection{Applications}
        Temporal networks are present in many areas of research interest, ranging from symptom interactions in mental health\cite{jordan2020current,contreras2020temporal}, to epidemiology\cite{masuda2013predicting}, protein interactions\cite{lucas2021inferring,jin2009identifying}, and social networks\cite{moinet2015burstiness,hanneke2010discrete}.

        % Overview of modelling attempts
        Generally attempts to model temporal networks either model the change in node state or the change in network structure.

        The node state describes some information about that node:
        \begin{align}
            _t x &= (m_1,\ldots, m_p)
        \end{align}
        Where $m_i$ is some metadata about the node $_t x$.

        An example of modelling the change in a node state could be the severity of a symptom at a given time, as in \Citeauthor{contreras2020temporal}\cite{contreras2020temporal}. Where the node $_t x$ represents a given symptom and $m_1$ would represent the severity of that symptom.

        % Intro contreras as attempt to model changes in nodes
        \Citeauthor{contreras2020temporal}\cite{contreras2020temporal} model the change in node state over time while the structure of the network remains constant. The authors look at the severity of symptoms over time, i.e. the state of the nodes, where the nodes exist in a statically structured temporal network.

        % Contreras methods
        \Citeauthor{contreras2020temporal} use multilevel vector autoregression, a method that finds correlations between the current state of a variable and the states of variables at previous time steps \cite{singer2003applied}, to model how the current symptoms of people with paranoia might predict their future symptoms. These models were then used to create three networks that linked symptoms.

        % Focal point of the paper is their temporal network
        Of particular interest to this thesis is the temporal network created by having a fully connected network of all symptoms including self-loops, where a symptom connects to itself; this self loop comes about from a symptom being correlated with itself from one time step to the next. The edge weights represent the extent to which the severity of each symptom at time $(t)$ predicts the severity of itself and other symptoms at time $(t+1)$.

@@

        % intro to paper as attempt to model node state
        In \cite{KARIMI20133476}\cite{KARIMI20133476}, the \Citeauthor{KARIMI20133476} look to model the change in node state while the network structure changes.

@@

        % what state the paper attempts to model
        The authors use a model to predict when a node will change from state 0 to state 1, in practice this could represent whether a user of a social network believes a rumour, or whether an ecological patch has been colonised.

@@

        % Methods to model node state
        To model the change in a nodes' state, the model uses a sliding time window and if the fraction of interactions with nodes of state 1 within that time window exceeds a given threshold, the node switches to state 1 as well.

        % Limitation of model
        In this work we aim to present a model for predicting the state of a node in the future. Instead, \Citeauthor{KARIMI20133476} aim to present a tool for understanding the spread of ideas, viruses, etc. within a network.

        Other efforts focus on modelling the changing structure of the network as in \Citeauthor{sanna2021link}\cite{sanna2021link}, which looks to model $_t G$ as a whole.

        % temp net representation
        In this case the \Citeauthor{sanna2021link} represent the temporal network as a series of adjacency matrices, each matrix representing an observation of the changing network.

        % Methods for embedding
        The authors look to model the structure of edges by using a spectral embedding (SVD). A spectral embedding uses the eigenvectors of a network's adjacency matrix to find vectors that represent the nodes of the network. The spectral embedding transforms their temporal sequence into a sequence of observations of a latent space.

@@

    \subsection{UDEs and NNDEs}
        % Justification for DEs being useful
        In some temporal networks the state of a node might influence the evolution of its neighbours. For example, in social networks we often observe that the more neighbours a node has, the more likely they are to gain more neighbours, a power law distribution of edges\cite{zhao2012multi,garg2009evolution}. We also see that the state of symptoms (occurrence, severity, and distress), can influence the states of other symptoms in patients undergoing chemotherapy \cite{papachristou2019network,kalantari2022network}, and for mental health disorders \cite{contreras2020temporal}. As well as in ecological networks where the interactions of species are often represented as a network, and the populations of one species may influence the population growth of another \cite{elton2001animal,volterra1927variazioni}; that is, the state (in this case the population) of a species (node), may influence the movement of another species.

@@

        % Can use a new NN structure to make this framework very general
        Our aim for this framework to be as general as possible and note that there may not yet be enough domain knowledge to perform traditional differential equation modelling to the desired accuracy. For example in network ecology models can be useful, but not perfect. Hence, we might wish to build on these models rather than trying to create something entirely new. This can be achieved by using this partial domain knowledge using it to create a UDE\cite{SciML_C_Rak}.

        \subsubsection{Theory}

            % Overview of why UDEs are useful
            Developed by \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, UDEs are a novel neural network architecture that aims to take the best of both traditional neural differential equation modelling and of flexible machine learning approaches to modelling.

@@

            % Software used for coding
            There is an ecosystem of packages specifically designed for performing machine learning with UDEs. The ecosystem is called Scientific Machine Learning (SciML)\cite{SciML_C_Rak}. Largely, the SciML ecosystem has been optimised for flexibility and efficiency with respect to models available and training performance. Given there does not seem to be any other ecosystem with this feature, it seemed like 
            an obvious choice.

        \subsubsection{Applications}
            Whilst there has been much interest in implementing UDEs and NNDEs, much of this work has been focussed on physics informed neural network and physics informed neural ordinary differential equations (PINNs and PINODEs) \cite{karniadakis2021physics,GAO2021110079,krishnapriyan2021characterizing,roehrl2020modeling}, and on improving modelling of fluids \cite{mahmoudabadbozchelou2021data,nguyen2022physics}. Alongside this, UDEs have gained popularity in the last few years, and there have been a number of studies exploring their usefulness in modeling the effect of restrictions due to COVID-19 on the virus' spread \cite{Dandekar2020.04.03.20052084}. Although there has been a large amount of research into the usefulness of these types of models, to the best of this author's knowledge they have never been applied to the problem of predicting temporal networks. This seems to be a natural fit for UDEs and NNDEs, given we may know relatively little about the processes that govern the evolution of temporal networks, and any information we do have can be incorporated into the model.

    \subsection{Symbolic Regression}
        % What is symbolic regression
        Symbolic regression is a type of analysis that looks to take input variables and find a relationship between them that describes the output variables. For symbolic regression, we construct a set of mathematical operations that may be used to find this relationship. The algorithm then generates combinations of these equations to find an expression that matches the output variables. Because we construct the set of potential operations, the resulting expression is interpretable by humans (a functional equation), as opposed to a NN for example (a numeric equation).

@@

        % Where has symbolic regression been used 
        Various methods of symbolic regression have been use in manufacturing systems, chemical systems, and tumor research\cite{can2011comparison,keith2021combining,yoshihara2013inferring}. As well research into improvements is still on going.

@@

        % Examples of use of genetic programming
        For example, both \Citeauthor{schmidt2009distilling,bongard2007automated} test their methods' capacity for obtaining a functional equation with physical simulation data, various pendula specifically. A ``meaningful'' equation in this case may be an equation that conserves momentum.

        % Expalaination of schmidt
        Examples of genetic programming include discovering equations governing the movement of simple harmonic and chaotic double pendula from data\cite{schmidt2009distilling}. In \Citeauthor{schmidt2009distilling}, the authors present an evolutionary algorithm to generate multiple candidate functions that approximate the partial derivatives of each variable that are calculated directly from the data.

@@

        % Benefits of combining UDEs and symbolic regression
        In very recent work, Kidger\cite{kidger2022neural} demonstrates that, with UDEs, two of the assumptions necessary for symbolic regression can be overcome. Those are the requirement for paired observations and derivatives, as well as the assumption that the function can be expressed as a shallow tree of symbolic operations. We are modelling a differential equation, and so will have access to the derivatives of our numerical model at any point, as neural networks are easily differentiable.

        % Usefulness of symbolic regression for nnde
        In \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, the authors demonstrate that by using symbolic regression to find a functional equation of our model, as opposed to just a UDE, we can improve the accuracy of extrapolating.
\end{section}

\begin{section}{Methodology}

        % Goal is to predict next time steps
        We are given a sequence of graphs, and our goal is to predict the next graph in the sequence.

        % How we achieve this goal
        To achieve this, we reinterpret our temporal network as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space.
        We then train a Neural Network Differential Equation (NNDE) to approximate the rate of change of the points in this latent space. With this numerical equation we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a temporal network.

        \begin{figure}
            \centering
            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology1.tex}}
                    \label{method net, a}
                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology2.tex}}
                    \label{method net, b}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology3.tex}}
                    \label{method net, c}
                \end{subfigure}

@@

                \end{tabular}
                \caption{Sequence of sparse adjacency matrices associated with the networks in \autoref{method nets}.}
                \label{method adjacency}
            \end{subfigure}

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                    $\begin{bmatrix}
                        -0.82  & 0.01\\
                        -0.64  & 0.36\\
                        -0.95 & -0.76\\
                        -0.64 &  0.36\\
                        -0.48 &  0.82\\
                        -0.17 & -0.40\\
                        -0.20 & -0.52\\
                        -0.07 &  0.25\\
                    \end{bmatrix}$

@@

                    $\begin{bmatrix}
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                         0.08 &  -0.69\\
                         0.11 &  -0.88\\
                         0.11 &  -0.88\\
                         0.08 &  -0.69\\
                    \end{bmatrix}$
                \end{tabular}

@@

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 1.png}
                        \label{}
                    \end{subfigure}

                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 2.png}
                        \label{}
                    \end{subfigure}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 3.png}
                        \label{}
                    \end{subfigure}
                \end{subfigure}

@@

    \subsection{Singular Value Decomposition and Random Dot Product Graphs}
            \label{svd}
            % Over view of section
            In this section we discuss the Singular Value Decomposition (SVD) of a matrix, and its usefulness in creating Random Dot Product Graphs (RDPG).

>>>>

\documentclass[12pt]{amsbook}
\usepackage[utf8]{inputenc}
\usepackage[
backend=biber,
style=numeric,
maxbibnames=99,
sorting=ynt
]{biblatex}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{svg}
\usepackage{gensymb}
\usepackage{float}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx, animate}
\usepackage{caption, subcaption}
\usepackage{pgffor}
\usepackage{mathtools}
\usepackage{array}
\usepackage{cleveref}
\addbibresource{bibliography.bib}
\usepackage{hyperref,thmtools}
%\usepackage{chngcounter}

\usetikzlibrary{decorations.markings}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta,positioning}

\counterwithin{figure}{chapter}
\counterwithin{section}{chapter}
\counterwithin{table}{chapter}
\counterwithin{subsection}{section}

\linespread{1.5}
\let\cleardoublepage\clearpage
% Override ugly default link
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = blue    %Colour of citations
}

@@

\newgeometry{left=3cm, right=3cm}
\begin{titlepage}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{uc-logo.png}\\
        \vspace{0.5cm}
        \begin{large}
        \textsc{School of Mathematics and Statistics\\ University of Canterbury \\
        Christchurch, New Zealand}\\
        \end{large}
        \vspace{0.2cm}
        \noindent\rule{\textwidth}{0.4pt}\\
        \begin{LARGE}
        \textbf{Modelling Temporal Networks with Scientific Machine Learning}\\
        \noindent\rule{\textwidth}{0.4pt}
        \end{LARGE}\\
        \vspace{0.2cm}
        \begin{large} 
        \textit{Submitted in partial fulfilment of the requirements for}\\
        \end{large}
        \begin{Large}
        \vspace{0.2cm}
        \textsc{The Degree of Master of Mathematical Sciences \\ in Computational and Applied Mathematics}\\
        \end{Large}
        \begin{large}
        \textit{by}\\
        \end{large}
        \begin{Large}
        \textsc{Connor Smith}\\
        \end{Large}
        \vfill
        \begin{large}
        \vspace{0.2cm}
        2023
        \end{large}
    \end{center}
\end{titlepage}
\restoregeometry
\pagestyle{plain}
{\renewcommand\addcontentsline[3]{}
\section*{Abstract}
}
    Temporal networks appear in applications ranging from symptom prediction to social networks to epidemiology. This thesis proposes a flexible framework for modelling temporal network progression which utilises a combination of singular value decomposition, scientific machine learning, and symbolic regression. We demonstrate the usefulness of the framework for modelling the progression of a variety of small temporal networks. Additionally, we present a Julia programming package for streamlining work with temporal networks. This thesis demonstrates our framework as a proof of concept, and we are confident that with further research, this framework will be a useful tool for modelling temporal networks in many fields.

\newpage
{\renewcommand\addcontentsline[3]{}
\section*{Acknowledgments}
}
    I'd like to thank my parents for teaching me to be curious, to enjoy my life at university, and to always strive to grow and learn. Without their love and support, I would not have made it to the point I am today.

    I'd like to thank my supervisors, Giulio and Miguel, who dedicated their time to help me overcome the countless challenges I faced over the course of this Thesis. Thanks to them, I am a better writer and researcher than I was when I started.

    Thanks to my office mates, who kept morale high by convincing me that every day was a good day for a pie. Without them, I would likely have given up a long time ago.

    Thanks to Trinity, whose love, support, and encouragement has been unwavering since day one.

{\hypersetup{hidelinks}\tableofcontents}

\pagestyle{headings}

\begin{chapter}{Introduction}
    Temporal networks are networks whose nodes and edges change in time. We aim, in this thesis, to describe the changing temporal network using mathematical language; that is to construct a model for the temporal network. 
    % Why temp nets are important
    Temporal network modelling is an important task in many real world applications; these include symptom interactions for mental health \cite{jordan2020current,contreras2020temporal}, epidemiology \cite{masuda2013predicting}, and protein interactions \cite{lucas2021inferring,jin2009identifying}.

    % Temp nets are dynam systs
    Temporal networks are constructed with nodes and edges. The state of a given node can be described by the set of edges connecting to it. As these edges change in time, so too does the state of the node. Hence, we can see temporal networks as a system of nodes whose states vary in time, a dynamical system.

    % Temp net events are discrete
    In the common representation of networks as binary-valued adjacency matrices (matrices where the observation or lack thereof, of an edge is represented as a 1 or 0 respectively) the events recorded in a temporal sequence of networks correspond to the appearance or the disappearance of links.

    % DEs are useful for modelling interacting states
    Differential equations are particularly useful for modelling systems where the state of one variable can effect the trajectories of other variables. We observe this behavior in temporal networks; nodes' connections within the network can influence the appearance of edges between other nodes, for example the phenomenon observed in \cite{newman2001clustering,capocci2006preferential}, where a node is more likely to gain the connections in the future the more connections it currently has (preferential attachment). With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling to predict the behaviour of a temporal network system using some initial state, i.e. the current observation of the network.

    % Difficult to model discrete events
    Because of the discrete nature of events, directly modelling the temporal networks as differential equations would require us to handle discrete jumps. The discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

@@

    % Our solution is interpretable because we use symb reg
    We then define and use Neural Network Differential Equations (NNDE)\cite{SciML_C_Rak} to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow us to bettor predict forward in time the evolution of the temporal networks.

@@

    % The framework can be applied to lots of fields
    As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations (UDEs).
\end{chapter}

\begin{chapter}{Literature Review}

    \section{Temporal Networks}
        % Use of temp nets
        \subsection{Definitions}
            Temporal networks are used to describe a series of interactions across time.
            A temporal network $G$ is defined as:  
            \begin{align}
                G=\{ {_t} G=({_t} N, {_t} E)\} _{t \in 1\ldots T } 
            \end{align}
            $_t x \in {_t} N, {_t} e\in {_t} E$, where ${_t} x, {_t} e$ are nodes and edges at time $t$ respectively, and ${_t} N, {_t} E$ are the sets of nodes and edges at time $t$ respectively.
            An example of this can be found in \cref{method nets}.
            \begin{figure}[H]
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology1.tex}}
                    \label{method net, a}
                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology2.tex}}
                    \label{method net, b}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.3\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology3.tex}}
                    \label{method net, c}
                \end{subfigure}

            \end{figure}
            %representations of temp nets as adj lists
            This can be represented in many ways, for example as a series of interactions. The series of interactions is generally represented as a list of tuples containing the nodes which interacted and the time that the interaction took place. 
            \begin{align}
                G=\{(i, j, t)|\space i,j\in {_t} N, (i\leftrightarrow j)\in {_t}E \}
            \end{align}

@@

        \subsection{Applications}
            Temporal networks are present in many areas of research interest, ranging from symptom interactions in mental health\cite{jordan2020current,contreras2020temporal}, to epidemiology\cite{masuda2013predicting}, protein interactions\cite{lucas2021inferring,jin2009identifying}, and social networks\cite{moinet2015burstiness,hanneke2010discrete}.

            % Overview of modelling attempts
            Generally attempts to model temporal networks either model the change in node metadata or the change in network structure.

            The node metadata describes some information about that node:
            \begin{align}
                _t x &= (m_1,\ldots, m_p)
            \end{align}
            Where $m_i$ is some metadata about the node $_t x$.

        \subsubsection{Modelling Node Metadata}
            An example of modelling the change in a node metadata could be the severity of a symptom at a given time, as in \Citeauthor{contreras2020temporal}\cite{contreras2020temporal}. Where the node $_t x$ represents a given symptom and $m_1$ would represent the severity of that symptom.

            % Intro contreras as attempt to model changes in nodes
            \Citeauthor{contreras2020temporal}\cite{contreras2020temporal} model the change in node metadata over time while the structure of the network remains constant. The authors look at the severity of symptoms over time, i.e. the metadata of the nodes, where the nodes exist in a statically structured temporal network.

            % Contreras methods
            \Citeauthor{contreras2020temporal} use multilevel vector autoregression, a method that finds correlations between the current value of a variable and the values of variables at previous time steps \cite{singer2003applied}, to model how the current symptoms of people with paranoia might predict their future symptoms. These models were then used to create three networks that linked symptoms.

            % Focal point of the paper is their temporal network
            Of particular interest to this thesis is the temporal network created by having a fully connected network of all symptoms including self-loops, where a symptom connects to itself (this self loop comes about from a symptom being correlated with itself from one time step to the next). The edge weights represent the extent to which the severity of each symptom at time $(t)$ predicts the severity of itself and other symptoms at time $(t+1)$.

@@

            % intro to paper as attempt to model node state
            In \Citeauthor{KARIMI20133476}\cite{KARIMI20133476}, \Citeauthor{KARIMI20133476} look to model the change in node metadata while the network structure changes.

@@

            % what state the paper attempts to model
            The authors use a model to predict when a node will change from value 0 to value 1, in practice this could represent whether a user of a social network believes a rumour, or whether an ecological patch has been colonised.

            % Methods to model node state
            To model the change in a nodes' metadata, the model uses a sliding time window and if the fraction of interactions with nodes of value 1 within that time window exceeds a given threshold, the node switches to value 1 as well.

@@

        \subsubsection{Modelling Network Structure}
            In this thesis, our focus is on modelling the changing structure of the temporal network as opposed to the changing node metadata.

            Efforts which focus on modelling the changing structure of the network include \Citeauthor{sanna2021link}\cite{sanna2021link}, which looks to model the structure of the temporal network as a whole.

            % temp net representation
            In this case \Citeauthor{sanna2021link} represent the temporal network as a series of adjacency matrices, each matrix representing an observation of the changing network.

            % Methods for embedding
            The authors look to model the structure of edges by using a Singular Value Decomposition (SVD). An SVD uses the eigenvectors of a network's adjacency matrix to find vectors that represent the nodes of the network. The SVD transforms their temporal sequence into a sequence of observations of a latent space.

@@

    \begin{section}{UDEs and NNDEs}
        % Justification for DEs being useful
        In some temporal networks the state of a node might influence the evolution of its neighbours. For example, in social networks we often observe that the more neighbours a node has, the more likely they are to gain more neighbours, a power law distribution of edges\cite{zhao2012multi,garg2009evolution}. We also see that the state of symptoms (occurrence, severity, and distress), can influence the states of other symptoms in patients undergoing chemotherapy \cite{papachristou2019network,kalantari2022network}, and for mental health disorders \cite{contreras2020temporal}. As well as in ecological networks where the interactions of species are often represented as a network, and the populations of one species may influence the population growth of another \cite{elton2001animal,volterra1927variazioni}; that is, the state (in this case the population) of a species (node), may influence the movement of another species.

@@

        % Can use a new NN structure to make this framework very general
        Our aim for this framework is to be as general as possible and note that there may not yet be enough domain knowledge to perform traditional differential equation modelling to the desired accuracy. For example in network ecology models can be useful, but not perfect. Hence, we might wish to build on these models rather than trying to create something entirely new. This can be achieved by using this partial domain knowledge using it to create a UDE\cite{SciML_C_Rak}.

        \subsection{Theory}

            % Overview of why UDEs are useful
            Developed by \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, UDEs are a novel neural network architecture that aims to take the best of both traditional differential equation modelling and of flexible machine learning approaches to modelling.

@@

            % Software used for coding
            There is an ecosystem of packages specifically designed for performing machine learning with UDEs. The ecosystem is called Scientific Machine Learning (SciML)\cite{SciML_C_Rak}. Largely, the SciML ecosystem has been optimised for flexibility and efficiency with respect to models available and training performance for UDEs. Given there does not seem to be any other ecosystem with this feature, it seemed like 
            an obvious choice.

        \subsection{Applications}
            Whilst there has been much interest in implementing UDEs and NNDEs, much of this work has been focussed on Physics Informed Neural Network, and Physics Informed Neural Ordinary Differential Equations (PINNs and PINODEs) \cite{karniadakis2021physics,GAO2021110079,krishnapriyan2021characterizing,roehrl2020modeling}, and on improving modelling of fluids \cite{mahmoudabadbozchelou2021data,nguyen2022physics}. Alongside this, UDEs have gained popularity in the last few years, and there have been a number of studies exploring their usefulness in modeling the effect of restrictions due to COVID-19 on the virus' spread \cite{Dandekar2020.04.03.20052084}. Although there has been a large amount of research into the usefulness of these types of models, to the best of this author's knowledge they have never been applied to the problem of predicting temporal networks. This seems to be a natural fit for UDEs and NNDEs, given we may know relatively little about the processes that govern the evolution of temporal networks, and any information we do have can be incorporated into the model.
    \end{section}
    \begin{section}{Symbolic Regression}
        % What is symbolic regression
        Symbolic regression is a type of analysis that looks to take input variables and find a relationship between them that describes the output variables. For symbolic regression, we construct a set of mathematical operations that may be used to find this relationship. The algorithm then generates combinations of these equations to find an expression that matches the output variables. Because we construct the set of potential operations, the resulting expression is interpretable by humans (a functional equation), as opposed to a NN for example (a numeric equation).

@@

        % Where has symbolic regression been used 
        Various methods of symbolic regression have been use in manufacturing systems, chemical systems, and tumor research\cite{can2011comparison,keith2021combining,yoshihara2013inferring}, and research into improvements is still on going.

@@

        % Examples of use of genetic programming
        For example, both \Citeauthor{schmidt2009distilling}\cite{schmidt2009distilling}, and \Citeauthor{bongard2007automated}\cite{bongard2007automated} test their methods' capacity for obtaining a functional equation with physical simulation data, various pendula specifically. A ``meaningful'' equation in this case may be an equation that conserves momentum.

        % Expalaination of schmidt
        Examples of genetic programming include discovering equations governing the movement of simple harmonic and chaotic double pendula from data. In \Citeauthor{schmidt2009distilling}\cite{schmidt2009distilling}, the authors present an evolutionary algorithm to generate multiple candidate functions that approximate the partial derivatives of each variable that are calculated directly from the data.

@@

        % Benefits of combining UDEs and symbolic regression
        In very recent work, Kidger\cite{kidger2022neural} demonstrates that, with UDEs, two of the assumptions necessary for symbolic regression can be overcome. Those are, the requirement for paired observations and derivatives, as well as the assumption that the function can be expressed as a shallow tree of symbolic operations. We are modelling a differential equation, and so will have access to the derivatives of our numerical model at any point, as NNs are easily differentiable.

        % Usefulness of symbolic regression for nnde
        In \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, the authors demonstrate that by using symbolic regression to find a functional equation of our model, as opposed to just a UDE, we can improve the accuracy of extrapolating.
    \end{section}
\end{chapter}

\begin{chapter}{Methodology}

        % Goal is to predict next time steps
        We are given a sequence of graphs, and our goal is to predict the next graph in the sequence. To achieve this, we reinterpret our temporal network as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space.
        We then train a Neural Network Differential Equation (NNDE) to approximate the rate of change of the points in this latent space. With this numerical equation, we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a temporal network.

        \begin{figure}[H]
            \centering

@@

                \end{tabular}
                \caption{Sequence of sparse adjacency matrices associated with the networks in \cref{method nets}.}
                \label{method adjacency}
            \end{subfigure}

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{m{0.03\textwidth} m{0.3\textwidth} m{0.26\textwidth} m{0.1\textwidth} l}
                &
                    $\begin{bmatrix}
                        -0.82  & 0.01\\
                        -0.64  & 0.36\\
                        -0.95 & -0.76\\
                        -0.64 &  0.36\\
                        -0.48 &  0.82\\
                        -0.17 & -0.40\\
                        -0.20 & -0.52\\
                        -0.07 &  0.25\\
                    \end{bmatrix}$

@@

                    $\begin{bmatrix}
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                         0.08 &  -0.69\\
                         0.11 &  -0.88\\
                         0.11 &  -0.88\\
                         0.08 &  -0.69\\
                    \end{bmatrix}$
                \end{tabular}
                \caption{Sequence of truncated SVDs associated with the networks in \cref{method nets}.}
                \label{method embed}
            \end{subfigure}

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 1.png}

@@

                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 2.png}

                    \end{subfigure}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 3.png}

                    \end{subfigure}
                \end{subfigure}

                \end{tabular}
                \caption{Sequence of embedded points associated with the networks in \cref{method nets}.}
                \label{method plots}

@@

    \begin{section}{Singular Value Decomposition and Random Dot Product Graphs}
        \label{svd}
        % Over view of section
        In this section we discuss the Singular Value Decomposition (SVD) of a matrix, and its usefulness in creating Random Dot Product Graphs (RDPG).
